---
title: "ADT Final Project Part 2 - R Markdown Code"
output:
  html_document: default
  pdf_document: default
date: "2023-04-01"
---

Team 11: Ameya Parab & Trishna Patil

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Importing and loading necessary libraries

```{r, eval = FALSE}
#install.packages("RMySQL")
#install.packages("DBI")

library(RMySQL)
library(DBI)
```


Connecting to DB

```{r, eval = FALSE}
con <- dbConnect(MySQL(), user = "root", password = "root1234", dbname = "doctors_appointment_db", host = "34.170.21.186", bulk = TRUE, prefer_large_objects = TRUE)

dbListTables(con)
```

Importing the raw data from CSV
Author: Ameya Parab

```{r, eval = FALSE}
file_path <- "C:\\Users\\ameya\\Documents\\IUB\\Spring 2023\\Applied Database Technologies\\Project\\DAC_NationalDownloadableFile.csv"
df <- read.csv(file_path, header = TRUE)

#filtering data to only contain data for state of Indiana and removing rows having null values for certain columns
filtered_df <- subset(df, st == "IN" & Med_sch != "OTHER" & Cred != "")
nrow(filtered_df)

write.csv(filtered_df, file = "C:\\Users\\ameya\\Documents\\IUB\\Spring 2023\\Applied Database Technologies\\Project\\filtered_df.csv", row.names = FALSE)

```

Inserting data into the different normalized tables 
Reference: https://www.geeksforgeeks.org/how-to-write-entire-dataframe-into-mysql-table-in-r/

STEP 3: Insert data in Doctors_temp table 
Author: Trishna Patil

```{r, eval = FALSE}
doctors_df <- filtered_df[, c("NPI", "frst_nm", "mid_nm", "lst_nm", "gndr")]
doctors_df <- doctors_df[!duplicated(doctors_df$NPI), ]
dbWriteTable(con, "Doctors_temp", doctors_df, overWrite=FALSE, append=TRUE)
```

STEP 5: Inserting into Contacts table
Author: Trishna Patil

```{r, eval = FALSE}
contacts_df <- filtered_df[, c("NPI", "phn_numbr")]
contacts_df$phn_numbr <- paste("(", substr(contacts_df$phn_numbr,0,3), ") -", substr(contacts_df$phn_numbr,4,6), "-",  substr(contacts_df$phn_numbr,7,10))

query <- paste0("INSERT INTO Contacts VALUES ", paste0("(", contacts_df$NPI, ", '", contacts_df$phn_numbr, "')", collapse = ", "))
dbExecute(con, query)
```

STEP 6: Inserting into Services table
Author: Trishna Patil

```{r, eval = FALSE}
services_df <- filtered_df[, c("NPI", "Telehlth", "ind_assgn", "grp_assgn")]
services_df <- services_df[!duplicated(services_df$NPI), ]

query <- paste0("INSERT INTO Services VALUES ", paste0("(", services_df$NPI, ", '", services_df$Telehlth, "', '", services_df$ind_assgn, "', '", services_df$grp_assgn, "')", collapse = ", "))
dbExecute(con, query)
```

STEP 7: Inserting into Schools table
Author: Trishna Patil

```{r, eval = FALSE}
schools <- filtered_df$Med_sch
schools <- unique(schools)

query <- paste0("INSERT INTO Schools (SchoolName) VALUES ", paste0("('", schools, "')", collapse = ", "))
dbExecute(con, query)
```


STEP 8: Inserting into Education table
Author: Trishna Patil

```{r, eval = FALSE}
education_df <- filtered_df[, c("NPI", "Med_sch", "Cred", "Grd_yr")]
dbWriteTable(con, "Education_temp", education_df, APPEND = TRUE)

```


STEP 9: Inserting into Addresses table
Author: Ameya Parab

```{r, eval = FALSE}
addresses_df <- filtered_df[, c("adrs_id", "adr_ln_1", "adr_ln_2", "cty", "st", "zip")]
addresses_df$zip <- paste(substr(addresses_df$zip,0,5), "-",substr(addresses_df$zip,6,9))
addresses_df <- addresses_df[!duplicated(addresses_df$adrs_id), ]

query <- paste0("INSERT INTO Addresses VALUES ", paste0("('", addresses_df$adrs_id, "', '", addresses_df$adr_ln_1, "', '", addresses_df$adr_ln_2, "', '", addresses_df$cty, "', '", addresses_df$st, "', '", addresses_df$zip, "')", collapse = ", "))
dbExecute(con, query)
```

STEP 10: Insert data in Organizations_temp table 
Author: Ameya Parab

```{r, eval = FALSE}
organizations_df <- filtered_df[, c("org_pac_id", "org_nm", "adrs_id", "num_org_mem")]
organizations_df <- organizations_df[!is.na(organizations_df$org_pac_id), ]
organizations_df <- subset(organizations_df, !duplicated(organizations_df$org_pac_id))
dbWriteTable(con, "Organizations_temp", organizations_df, overWrite=FALSE, append=TRUE)
```

STEP 12: Inserting into DoctorClinics table
Author: Ameya Parab

```{r, eval = FALSE}
clinics_df <- filtered_df[, c("NPI", "adrs_id", "org_pac_id")]
clinics_df <- clinics_df[is.na(clinics_df$org_pac_id), ]
clinics_df <- clinics_df[!duplicated(clinics_df$NPI), ]

query <- paste0("INSERT INTO DoctorClinics VALUES ", paste0("(", clinics_df$NPI, ", '", clinics_df$adrs_id, "')", collapse = ", "))
dbExecute(con, query)
```

STEP 13: Inserting into DoctorOrganizations table
Author: Ameya Parab

```{r, eval = FALSE}
doctor_organizations_df <- filtered_df[, c("NPI", "org_pac_id", "adrs_id")]
doctor_organizations_df <- doctor_organizations_df[complete.cases(doctor_organizations_df$org_pac_id), ]
duplicates_df <- doctor_organizations_df[!duplicated(doctor_organizations_df$adrs_id), ]
duplicates_df$IsHospital <- rep(1, nrow(duplicates_df))
uniques_df <- doctor_organizations_df[duplicated(doctor_organizations_df$adrs_id), ]

query <- paste0("INSERT INTO DoctorOrganizations VALUES ", paste0("(", duplicates_df$NPI, ", '", duplicates_df$org_pac_id, "', '", duplicates_df$IsHospital, "')", collapse = ", "))
dbExecute(con, query)
```

STEP 14: Inserting into Specialties table
Author: Trishna Patil

```{r, eval = FALSE}
specialties <- c(filtered_df$pri_spec, filtered_df$sec_spec_1, filtered_df$sec_spec_2, filtered_df$sec_spec_3, filtered_df$sec_spec_4)
specialties <- unique(specialties)

query <- paste0("INSERT INTO Specialties (SpecialtyName) VALUES ", paste0("('", specialties, "')", collapse = ", "))
dbExecute(con, query)
```

STEP 15: Inserting into DoctorSpecialties table
Author: Trishna Patil & Ameya Parab

```{r, eval = FALSE}
doctor_specialties_df <- filtered_df[, c("NPI", "pri_spec", "sec_spec_1", "sec_spec_2", "sec_spec_3", "sec_spec_4")]

primary_specialty_df <- data.frame(NPI = doctor_specialties_df$NPI, specialty = doctor_specialties_df$pri_spec)
primary_specialty_df$IsPrimary <- rep(1, nrow(primary_specialty_df))


secondary_specialty1_df <- data.frame(NPI = doctor_specialties_df$NPI, specialty = doctor_specialties_df$sec_spec_1)
secondary_specialty1_df[complete.cases(secondary_specialty1_df$specialty), ]
secondary_specialty1_df$IsPrimary <- rep(0, nrow(secondary_specialty1_df))

secondary_specialty2_df <- data.frame(NPI = doctor_specialties_df$NPI, specialty = doctor_specialties_df$sec_spec_2)
secondary_specialty2_df <- secondary_specialty2_df[complete.cases(secondary_specialty2_df$specialty), ]
secondary_specialty2_df$IsPrimary <- rep(0, nrow(secondary_specialty2_df))

secondary_specialty3_df <- data.frame(NPI = doctor_specialties_df$NPI, specialty = doctor_specialties_df$sec_spec_3)
secondary_specialty3_df <- secondary_specialty3_df[complete.cases(secondary_specialty3_df$specialty), ]
secondary_specialty3_df$IsPrimary <- rep(0, nrow(secondary_specialty3_df))

secondary_specialty4_df <- data.frame(NPI = doctor_specialties_df$NPI, specialty = doctor_specialties_df$sec_spec_4)
secondary_specialty4_df <- secondary_specialty4_df[complete.cases(secondary_specialty4_df$specialty), ]
secondary_specialty4_df$IsPrimary <- rep(0, nrow(secondary_specialty4_df))

nrow(primary_specialty_df)
combined_df <- rbind(primary_specialty_df, secondary_specialty1_df, secondary_specialty2_df, secondary_specialty3_df, secondary_specialty4_df)
nrow(combined_df)
dbWriteTable(con, "DoctorSpecialties_temp", combined_df, APPEND = TRUE)
```

```{r, eval = FALSE}
library(RColorBrewer)
# Create example data frame

credentials_df <- dbGetQuery(con, "SELECT Credential, COUNT(*) AS Num_Of_Doctors FROM Education WHERE Credential != '' GROUP BY Credential LIMIT 10")
color <- brewer.pal(length(credentials$Credential), "Set3")
# Create bar plot
plot_ly(credentials_df, x = ~Credential, y = ~Num_Of_Doctors, type = 'bar', colors = list(color)) %>%
layout(yaxis = list(type = 'log'))
```

```{r, eval = FALSE}
library(leaflet)
#install.packages("tmaptools")
library(tmaptools)
address_df <- dbGetQuery(con, "SELECT CONCAT(d.FirstName, ' ', d.LastName) AS DoctorName, CONCAT(a.AddressLine1, ', ', a.City, ', ', a.State) AS Address
                                      FROM Addresses a
                                      INNER JOIN DoctorClinics dc
                                      ON a.AddressID = dc.AddressID
                                      INNER JOIN Doctors d
                                      ON dc.DoctorId = d.DoctorId")

address_df <- address_df[complete.cases(address_df), ]
#coord_df <- data.frame(Name = character(0), lat = character(0), lng = character(0))
names <- list()
lat <- list()
lng <- list()

sum(is.na(address_df$Address))
for (i in 1:(length(address_df$Adress - 1))) {
     if(length(geocode_OSM(address_df$Address[i])) != 0)
      cod <- geocode_OSM(address_df$Address[i])
      names <- c(names, address_df$DoctorName[i])
      lat <- c(lat, cod$coords[1])
      lng <- c(lng, cod$coords[2])
}

coord_df <- data.frame(
  Name = names,
  lat = lat,
  lng = lng
)
#coord_df$lat <- coords_df$coords[1]
#coord_df$lng <- coords_df$coords[2]

coord_df
leaflet(coord_df) %>%
      addTiles() %>%
      addMarkers(label = ~DoctorName, popup = ~DoctorName, clusterOptions = markerClusterOptions())
```

```{r, eval = FALSE}

doctors_df <- dbGetQuery(con, "SELECT DISTINCT(DoctorID), DoctorName, SpecialtyName, Credential, GraduationYear, Teleconsultation FROM doctor_details")
doctors_df
```


```{r, eval = FALSE}

doctors <- dbGetQuery(con, "SELECT DISTINCT(DoctorID) FROM doctor_details LIMIT 10")
for(i in 1:length(doctors$DoctorID)){
  query <- paste0("INSERT INTO AvailableAppointments VALUES (", doctors$DoctorID[i], ",'2023-05-07', '9:00')")
  dbExecute(con, query)
}
```

```{r, eval = FALSE}
library(dplyr)
library(tidygeocoder)
dc_addresses <- tribble( ~name,~addr,
"White House", "1600 Pennsylvania Ave Washington, DC",
"National Academy of Sciences", "2101 Constitution Ave NW, Washington, DC 20418",
"Department of Justice", "950 Pennsylvania Ave NW, Washington, DC 20530",
"Supreme Court", "1 1st St NE, Washington, DC 20543",
"Washington Monument", "2 15th St NW, Washington, DC 20024")
coordinates <- dc_addresses %>%
  geocode(addr)
print(coordinates)
```
```{r, eval = FALSE}
address_df <- dbGetQuery(con, "SELECT DISTINCT DoctorId, DoctorName, Address 
FROM (SELECT d.DoctorId, CONCAT(d.FirstName, ' ', d.LastName) AS DoctorName, CONCAT(a.AddressLine1, ', ', a.City, ', ', a.State) AS Address
	FROM Addresses a
	INNER JOIN DoctorClinics dc
	ON a.AddressID = dc.AddressID
	INNER JOIN Doctors d
	ON dc.DoctorId = d.DoctorId
UNION 
	SELECT d.DoctorId, CONCAT(d.FirstName, ' ', d.LastName) AS DoctorName, CONCAT(a.AddressLine1, ', ', a.City, ', ', a.State) AS Address
	FROM Addresses a
	INNER JOIN Organizations org
	ON a.AddressID = org.AddressID
	INNER JOIN DoctorOrganizations dorg
	ON org.OrganizationID = dorg.OrganizationID
	INNER JOIN Doctors d
	ON dorg.DoctorId = d.DoctorId) AS doctor_addresses")

lat <- list()
long <- list()
for (i in 1:length(address_df$Address)) {
      addresses <- tribble( ~name, ~addr, address_df$DoctorName[i], address_df$Address[i])
      geocoded_results <- addresses %>% geocode(addr)
      lat <- c(lat,geocoded_results$lat)
      long <- c(long,geocoded_results$long)
}
    
coord_df <- address_df[, c("DoctorId", "DoctorName")]
coord_df$lat <- unlist(lat, recursive = TRUE)
coord_df$lng <- unlist(long, recursive = TRUE)
dbWriteTable(con, "Coordinates_Temp", na.omit(coord_df), overWrite=FALSE, append=TRUE)
```

```{r, eval = FALSE}

```

Disconnect Database

```{r, eval = FALSE}
dbDisconnect(con)
```